---
title: "ECOL 597 class notes"
output: html_document
date: "last knit `r Sys.time()`"
---


#  2021-01-19

Chris, Lillian (Lilly)



#  2021-01-21

Sydney- NC, Rome

keyword library- lists the words you can't use in Python

use `input` to allow user input to assign an object later 

for logicals (booleans) in Python, case is important: True or False
 - true is 1, false is 0. Same as R. 

data types....float vs integer. Can't multiply integers? First have to convert to float? [**follow-up**: it didn't work because one of the varibales was saved as a string.]

adding strings together: this is like `paste()` or `stringr::str_c()` in R

and you can vectorize it by multiplying one or more strings by a number








#2021-01-26

Sam, Masie

conditionals and logical operators 

while loops : use a flag at the beginning 
example of asking user to input the answer to "what is 2+3?"

In the end what matters is that what follows the `while` must ONLY evaluate if the condition is satisfied 


`not` has a special meaning in Python; it is similar to !=

5 min break, talked to Jayden about programming languages 


for loops  vs while loops.....counter is incremented at end of the "action"

lists in Python are composed using square brackets 

for i in range(10) . do something. However range starts at 0 not 1. EXCEPT if you give it a start and stop number it will stop at the stop number. In other words, the behavior of `range` changes depending on the arguments you pass it. Best way to avoid problems is to explicitly state what the arguments are. 



# 2021-01-28

Why use functions ?

modularity, shareability, debugging

local vs global variables (I think this is called lexical scoping ?)

parameter vs argument. Semantically different so good to know this but in practice they are used interchangeably. 

as a good practice, try to never use a global variable inside a function. You might get behavior you weren't expecting and more importantly it is NOT PORTABLE. 


to allow a parameter to have an arbitrary number of arguments, use the *operator prior to the parameter when defining the function

- alternatively, the parameter can just be typed as is when defining the function, but then you have to pass a list as the argument. 

skpping ahead so we can cover random number generation

they are never truly random....has something to do with your system clock and a complicated formula that decides what to do next. Keep this in mind when running simulations. 

Ability to set random seed (glad to have discovered this in R)

differences library, module, and package are minute although real; many people use them interchangeably. 

When you refer to functions in python, you always use the library name and then a dot to precede the function. same as `pkg::function()` in R.

In Python, creating a module is as simple as saving a python file and then calling `import <filepath>. 




# 2021-02-02

- long asides on homework and individual questions 
- installed Anaconda
-eventually started discussing strings
  - indexing is pretty similar to R....a few important differenes 
- use a negative number to indiate you want to start at the end of the string when looking for a single character. Just like **stringr**
  - the colon operator can be used before OR after a number when defining a sequence, for example word[:-5] would give all characters from zero to the one 5 from the end. The starting the count at zero thing does not apply for negative numbers. 
 - the `in` operator is useful; it is similar to `%in%` in R. Returns a logical about whether a given string is detected inside another string
- to change case, type the stringa and  `.upper()`
  - this is the general way to apply a function to an existing object, i.e. appending an existing list.


# 2021-02-23

Starting R 

logicals - the vectorized nature of `&` c.f. `&&`

```{r}
a <- c(5, 10, 15, 20)
b <- c(2, 4, 16, 256)

# try it with slicing 

a[a > b & b < a]
a[a > b && b < a]
a[a > 5 & a < 20]


```

Take user input with `readline()`

```{r}

sort(sample(1:10), decreasing = T)
sample(1:10)
```

